\newcommand{\mybookblurb}{%

We live in an era where the rules are being rewritten before our eyes. When AI lifts everyone's abilities to an 80 out of 100, 80 becomes the new zero. This book sets out, in the plainest language I can manage, to work through three questions with you: How does AI actually work under the hood? When AI can do so many jobs better than we can, what is it that AI still \emph{cannot} do? And how do we harness AI to thrive in this new era---to become truly irreplaceable?

\par
The book begins with \emph{connection} and \emph{latency}, explaining how information enters our lives. It then confronts the central paradox of data: data looks as abundant as an oil field, yet data that cannot be used is not an asset---it is a liability. We should pursue data that is complete and deep, not merely ``big,'' and above all, we must catch the anomalies that vanish in a blink. Along the way, you will learn to see through the many tricks data can play, to understand how data can be put to work without being leaked, and to protect your privacy. Large language models look like magic, but the principle beneath is not mysterious at all: they simply predict the next word from the words that came before. What AI does best is produce the standard answer; what it handles worst is the rare and unusual. This is true for models, and it is true for people. The cheaper the standard answer becomes, the more precious it is to be different.

\par
This book is for everyone who feels both curious and anxious about the age of AI: whether you are a student like my daughter, or a parent like me worrying about your child's future; whether you run your own business or punch the clock for someone else; or whether you simply want to understand, ``What does AI actually have to do with my life?''

}

\newcommand{\mybookrights}{%
No part of this book may be reproduced or transmitted in any form or by any means without permission.\\
All rights reserved.}

\vspace* {-3em}

\newcommand{\mybookcipheading}{图书在版编目（CIP）数据}
\newcommand{\mybookcipbody}{%
\hspace*{2em}《\mytitlemain》 / \myauthorzh\ 著.\\
\hspace*{2em}北京：\mypublisher，2026. -- ISBN 待正式出版后补充\\
\hspace*{2em}I. 待补充\\
\hspace*{2em}中国国家版本馆 CIP 数据核字待补充}

\newcommand{\mybookpublicationinfo}{%
责任编辑：董 英\\
印\qquad 刷：待补充\\
装\qquad 订：待补充\\
出版发行：\mypublisher\\
\hspace*{4.9em}北京市海淀区万寿路173信箱\qquad 邮编：100036\\
开\qquad 本：待补充\qquad 印张：待补充\qquad 字数：待补充\\
版\qquad 次：待补充\\
印\qquad 次：待补充\\
定\qquad 价：待补充}

\newcommand{\mybookserviceinfo}{%
凡所购买电子工业出版社图书有缺损问题，请向购买书店调换。若书店售缺，请与本社发行部联系，联系及邮购电话：（010）88254888，88258888。\\
质量投诉请发邮件至 zlts@phei.com.cn，盗版侵权举报请发邮件至 dbqq@phei.com.cn。\\
本书咨询联系方式：faq@phei.com.cn。}

\mybookinfopage
