\guidepart{Prologue}{

This part introduces the magical world of the Fifth Dimension. It serves as the prologue to the book and as a panoramic overview. More precisely, it is a navigational chart for the age of AI.

\vspace{1em}

On your first reading, do not try to memorize every spell and its details. Just take in the silhouette: in this world made of bits, why does old experience gradually lose its power? The chapters that follow will unpack each spell introduced here, one by one.

}

\chapter*{The Magical World of the Fifth Dimension}
\phantomsection
\addcontentsline{toc}{chapter}{The Magical World of the Fifth Dimension}

\markboth{The Magical World of the Fifth Dimension}{}

\begin{myquote}
The best way to appreciate the value and impact of ``Being Digital'' is to reflect on the difference between bits and atoms.

\hfill ---\textit{Being Digital}

\end{myquote}

\section{First Intimate Contact}

As I typed the title ``First Intimate Contact,'' I could not resist revisiting Pizi Cai's legendary novel from 1998. Those were the wild frontier days of the internet. I was a college freshman, spending my days posting on the 88 Forum. There was a famous line from that era, now rarely mentioned: ``On the internet, nobody knows you're a dog.''\footnote{``On the Internet, nobody knows you're a dog.'' The caption of a cartoon by Peter Steiner, published in \textit{The New Yorker} on July 5, 1993.}

Just as God's creation unfolded---first dividing heaven and earth, then arranging the sun, moon, and stars, the birds and the fish, and only on the sixth day creating humankind---when we began building an entirely new world through the technology of bits, the first things we erected were portal sites (Sina, NetEase) and BBS chat rooms (Shuimu Tsinghua, Piaomiao Shuiyun) and other content infrastructure.

Back then, people navigated this world mostly through screen names, email addresses, and temporary identities, and the connection to the physical world was far looser than it is today. Around the year 2000, the bridge between this world of bits and real-world identity systems was still fragile, and business models wobbled accordingly.

\vspace{1em}

The dot-com bubble burst. \textbf{From heaven, back to hell.}

What eventually gave the internet a viable business model was a key technology: the Cookie ID plus click logs. Websites wrote local identifiers into browsers via cookies. These identifiers did not necessarily correspond to a real identity, but they were enough for a platform to string together browsing sessions and search click logs under a single tag.

\textbf{With} these identifiers and click logs, a direct link was forged between real, living human beings and the world built from bits.
For the first time, platforms could stitch discrete page visits into the behavioral arc of ``the same user'' with reasonable stability. As long as cookies were not cleared, a vast trail of behavior within a single site could be continuously recorded. Later, cross-site account systems, advertising IDs, device fingerprinting, and other technologies further strengthened this tracking capability.

\vspace{1em}

As we spent more and more of our lives on the internet, everything we did in the four-dimensional world began, step by step, to be projected into this brand-new world built from bits. And so, for the first time in this world of bits, \textbf{there were real people.}
{\kaiti (Some contrarian will point out that there is also a kind of magic---like the ancient Chinese sorcery of scattering beans to summon soldiers---that can conjure hordes of web-crawling robots, and those are decidedly not real people. We will discuss this particular magical skill in a later chapter.)}

\vspace{1em}

By analyzing these click logs, internet advertising companies gradually learned to guess our preferences and serve content and ads accordingly. Alibaba, Baidu, Facebook\footnote{The first draft of this prologue was written in 2020, when Meta was still called Facebook. Looking back, most of what was written in this chapter has stood the test of time, so I have largely kept the original phrasing and tone.}, Google---companies whose core business model is advertising---are all built on top of this capability. Does this sound familiar by now? Yes: this technology, carried to its present state, is what we know as precision advertising, and in certain scenarios it has evolved into the much-debated practice of ``big data price discrimination''---charging loyal customers more.

\vspace{1em}

Day after day, a version of ``us'' is being assembled\footnote{The more technical term for this is \textit{digital twin}.} from bits in this world---or more precisely, an ever-more-detailed data portrait. It does not truly possess all of our memories, yet it accumulates more and more traces of our behaviors, preferences, and habits. This bit-ified ``us'' comes to know, with increasing accuracy, which airlines we prefer, what food we eat, what movies we watch, and so on.

\begin{thinkbox}[Reflection]
The Ship of Theseus paradox: is this digitally assembled person still you?
\end{thinkbox}

\newpage
\begin{myquote}
The Ship of Theseus paradox originates in the writings of Plutarch.

It describes a ship that could sail the seas for hundreds of years, thanks to continuous repair and replacement of its parts. Whenever a plank rotted, it was swapped out, and so on, until every single component had been replaced.

\vspace{1em}

The question is: is the ship that ultimately remains still the original Ship of Theseus?

Or is it an entirely different ship?

If it is no longer the original, at what point did it cease to be?

\vspace{1em}

The philosopher Thomas Hobbes later extended the puzzle: if someone were to build a new ship from all the old parts removed from the Ship of Theseus, which of the two ships would be the real Ship of Theseus?
\end{myquote}

\section{Will a Dark Future Ruled by AI Ever Arrive?}

The world did not stop at being recorded.

\vspace{1em}

Recording was only the beginning.

When data becomes rich enough, connections broad enough, and algorithms mature enough,

recording keeps growing forward:

first comes explanation, then prediction;

first comes assisted judgment, then participation in decisions;

first comes speaking on our behalf, then acting in our place.

\vspace{1em}

And so that question we have posed again and again in science fiction suddenly feels much closer to reality.
If one day even explanation, judgment, and choice begin to be handed over, piece by piece, to machines, will we slowly be edged aside in the very world we built with our own hands?

Perhaps what truly unsettles us is not ``Will it rule over us?'' but rather: when it begins to understand for us, predict for us, decide for us---are we willing to let it?

\vspace{1em}

Let us not rush to answer.

Before we do, perhaps we need to see clearly what kind of place we are walking into.
\newpage
\section{The Bit-Journey of the Fifth Dimension: From Recording to Intelligence}

Let us shift our perspective on the question above. \textit{Being Digital} prophesied the migration from atoms to bits; today we have embarked on the next leg of the bit-journey: from recording to intelligence. This is not an entirely new world so much as a new dimension growing out of the four-dimensional world we know. In this book, I call this higher-dimensional world \textbf{the Fifth Dimension}.

\begin{myquote}
The magical powers of a three-dimensional being over a two-dimensional one:

In two-dimensional space, the shortest distance between two points is a straight line.

A being living in three-dimensional space can bend the fabric of a two-dimensional plane,

so that two distant points overlap into one.

This is an example of how a higher-dimensional being possesses superpowers in a lower-dimensional space.
\end{myquote}

The four-dimensional world we inhabit obeys the famous Second Law of Thermodynamics: all things tend from order toward disorder. The concept of ``entropy'' in physics was proposed to measure the degree of this disorder; Claude Shannon later introduced ``information entropy'' into the field of information theory. Unlike the four-dimensional world, which drifts inexorably toward chaos, the Fifth Dimension---powered by \textbf{algorithms}, an alchemy whose purpose is to reduce information entropy---will become a world of ever-increasing efficiency and ever-improving order.\footnote{Of course, this order comes at a cost: energy consumption in the four-dimensional world. And as AI turns energy into computing power, and computing power into labor, the share of energy as a factor of production will only keep growing!}

\vspace{1em}

We still live in the physical world, of course---still eating, sleeping, still bound by time and distance. But at the same time, our memories, relationships, expressions, transactions, and identities are being translated, bit by bit, into another space. It is not our whole flesh-and-blood selves that cross over; rather, another ``us,'' pieced together from data, begins to leave traces there, accumulating capabilities.

\vspace{1em}

Once inside this world, many things that are slow, cumbersome, and expensive in the physical world begin to happen in an entirely different way. Distance is compressed. Copying costs almost nothing. Choices can be amplified. And so, abilities that once looked like magic start becoming everyday.

% 这些能力会像哆啦A梦的魔法口袋一样，五彩缤纷又触手可及。

% 接下来，让我们开启这段第五维度的魔幻旅程。在这个全新的世界，从这个魔法口袋里掏出的第一样东西是什么——咫尺天涯。

Now, let us embark on this magical journey through the Fifth Dimension. Let us see what the first magical ability looks like in this ever-more-intelligent world.

\newpage
\subsection{Spell One: A World Apart---How Information Collapses Time and Space}

A World Apart is the first spell of the Fifth Dimension: it makes data transmissible.

\vspace{1em}

From the moment humanity sent its first telegraph, we held the key to the Fifth Dimension. Before that, the fastest way to transmit information was to shout. Sound travels through air at roughly 340~m/s, but its range does not exceed 100~meters (with the exception of Cecilia Cheung, who mastered the ``Lion's Roar'' technique). The more common method was the letter:

\begin{myquote}
红藕香残玉簟秋。轻解罗裳，独上兰舟。

云中谁寄锦书来？雁字回时，月满西楼。

花自飘零水自流。一种相思，两处闲愁。

此情无计可消除，才下眉头，却上心头。
\end{myquote}

This is Li Qingzhao's ``Yi Jian Mei''---one of my all-time favorite \textit{ci} poems.\footnote{Li Qingzhao (1084--c.\,1155) was one of China's greatest poets. ``Yi Jian Mei'' (literally ``A Twig of Plum Blossoms'') is a lyric about longing for a far-away beloved: ``Red lotus fragrance fades, the jade mat feels autumn's chill\ldots\ Who sends a brocade letter through the clouds? When the wild geese return in formation, the moon floods the western tower.'' The poem closes with an ache that ``no sooner leaves the brow than it steals back into the heart.''} Our children, born into the Fifth Dimension, can scarcely imagine the homesickness of those days spent craning one's neck for a letter.\footnote{Their generation is called ``digital natives,'' more accustomed to the perfunctory three-word check-in: ``Ate. Showered. Sleeping.''}

\vspace{1em}

We all know that one important measure of an asset is its turnover rate.

If we think of data as an asset, then the speed of data transmission is directly proportional to asset turnover:

the faster data travels, the higher the turnover.

\vspace{1em}

Before entering the Fifth Dimension, because information did not flow freely, the supply of information was severely inadequate.

In the Fifth Dimension, the teleportation of A World Apart operates at the speed of light.

\vspace{1em}

At the speed of thought, we can deliver a piece of information to a distant point in the Fifth Dimension within milliseconds. Space is no longer humanity's shackle. If we simply compare the speed of sound in air with the speed of electromagnetic signals, the difference in magnitude is close to a millionfold (roughly $3\times10^8$~m/s vs.\ 340~m/s).
It is precisely this orders-of-magnitude leap\footnote{This does not mean that real-world communication efficiency improves by nearly a millionfold, because encoding, devices, network congestion, and human response times still create friction.} that rewrote ``waiting for a message'' into ``real-time interaction,'' vastly boosting the efficiency of communication and collaboration in the Fifth Dimension.

\newpage
\textbf{In} our global village, we are already accustomed to instant messaging via WeChat and WhatsApp, and in our leisure time we can fire up a real-time strategy game like StarCraft. The next time we will need the kind of non-instant, ``gazing-at-the-wild-geese'' communication will be when the global village starts talking to the Mars village. In other words, once we leap from the four-dimensional world to the Fifth Dimension, all we need to do is launch a rocket to deliver a receiver to Mars and set it up; the Fifth Dimension will automatically extend itself there. When that day comes, the magical beings living in the Fifth Dimension can ride a lightspeed rocket to Mars Village. That will mark the dawn of humanity's interstellar colonial era.

\subsection{Spell Two: Shadow Clones---Complete Data Is the Real Big Data}

Shadow Clones is a ninjutsu from the anime \textit{Naruto}: by expending energy, you can replicate yourself into as many copies as you wish.
In quantum physics, however, there is a no-cloning theorem\footnote{Wootters and Zurek, \textit{A single quantum cannot be cloned}, Nature, 1982.}: an unknown quantum state cannot be perfectly copied. In schemes like quantum teleportation, the original state is destroyed precisely to avoid producing two identical copies of ``you.''

\vspace{1em}

In the four-dimensional world, means of production such as oil and minerals are gradually consumed as production advances. Shadow Clones, the second spell of the Fifth Dimension, enables near-lossless copying of this world's most important means of production---data.

\vspace{1em}

My university mentor, Professor Cao Jiashu, taught biological evolution. He had a maxim that has stayed with me my entire life: {\kaiti ``I do not ask you to be far ahead of everyone else---only to rise above them by the height of a single leaf. Achieve that, and all the sunlight and rain in the sky belong to you.''}


On further reflection, the lesson I drew from biological evolution was not only vertical---making yourself better, so that tomorrow's self surpasses yesterday's---but also horizontal: competing simultaneously against other individuals and other strategies. \textbf{Against the weak, you may be slightly weak; against the strong, you must be stronger still.}

``What others consume, you lose.'' Finite resources create competitive pressure; in biological evolution this is one of the most common---and most brutal---selection criteria.

\begin{importantbox}
Most competitive strategies are, in some sense, zero-sum games.
\end{importantbox}

\textbf{The} Shadow Clones spell does not conjure scarcity out of existence. What it does, for the first time in large-scale social collaboration, is introduce a critical factor of production that is not depleted by use---and may, in fact, grow richer the more it is used. Over the long run, this change will create conditions for more cooperative strategies, reshape the nature of competition, and push the world of the Fifth Dimension toward a win-win future.

Yet Shadow Clones is also a double-edged sword. The cost of copying data is so low that a single misstep---say, a loop that keeps replicating itself---can fill up all available storage. Data can metastasize like cancer cells, devouring every last byte of finite storage until healthy data starves and the host and the cancer perish together. People usually understand ``big data'' to mean ``the more data, the better.'' This chapter aims to explore big data from a fresh angle: \textbf{data completeness}.

\subsection{Spell Three: The Stealth Codex---How We Protect Our Digital Assets}
In science fiction movies, invisibility typically means vanishing from sight, slipping into shadow. But in the Fifth Dimension, true invisibility is no longer about making an object unseen. The Stealth Codex\footnote{The Stealth Codex (\textit{Dunjia Tianshu}) originally referred to the arts of concealment and evasion in China's ancient divinatory traditions. Here, it borrows the idea of ``hiding one's form and erasing one's tracks'' as a metaphor for the ability, in the digital world, to verify and compute information without exposing its raw content.} is the third spell of the Fifth Dimension: it allows data to be used without revealing its contents.

In the four-dimensional world, collaborative interaction is usually open: to prove you own something, you must show it; to prove you are over eighteen, you must present your ID card, exposing your name, address, and even your photograph.
In the flow of information, privacy is laid bare before the verifier.

In the Fifth Dimension, however, technologies such as zero-knowledge proofs make it possible to prove that you know something without disclosing any specifics about what you know. The essence of this spell is that one party (the prover) can demonstrate to another party (the verifier) that a given statement is true, without revealing any information beyond the truth of that statement.

\vspace{1em}

\textbf{What does this have to do with ordinary readers like us?}

If data is the core asset of the future yet can be replicated by Shadow Clones, how can we protect our assets while still putting them to use?
The greatest transformation brought by The Stealth Codex is that it lays the foundation for trusted computation. It resolves the long-standing core tension in the digital world---the balance between privacy and trust. In later chapters, we will introduce readers to fascinating algorithms such as the Millionaires' Problem, and explore how to build a transparent, auditable digital society while safeguarding privacy.


\subsection{Spell Four: The Moonlight Box---How to Predict the Future from the Past}
The fourth dimension of the four-dimensional world is time. Unlike the three spatial dimensions of length, width, and height---where movement is reversible---time flows in only one direction.

\begin{myquote}
An inch of time is worth an inch of gold,

yet an inch of gold cannot buy an inch of time.

Time is like river water---

it can only flow onward, never back.
\end{myquote}

This poem is plain and unadorned, yet it captures the essence of time as a dimension: it flows in one direction only. In the four-dimensional world, time travel exists only in fantasy films.\footnote{According to the principle of locality in relativity, information cannot travel faster than light; in the four-dimensional world, time travel is impossible.} The Moonlight Box\footnote{A magical artifact from Stephen Chow's film \textit{A Chinese Odyssey}; it can reverse time, sending its user back to the past in an attempt to change the future.} is the fourth spell of the Fifth Dimension. It lets us do what fantasy films do: predict the future from the past.

\vspace{0.8em}

Thanks to the second spell, Shadow Clones, we can replicate the world. Each time the cloning spell is cast, a marker called a \textit{timestamp} is attached, recording the moment of replication. Because timestamps exist, we can use \textbf{snapshot} technology to perform temporal rollbacks on any part---or all---of the Fifth Dimension, returning to any point in time. Readers who have played video games will find this familiar: it is saving and loading.

\vspace{0.8em}

In the Fifth Dimension, as the river of time flows forward, the system simultaneously creates save points. By recording only the small amount of data that changes at each point in time, it can describe the complete state at any moment. This technique was first introduced in data warehousing to describe the full state of any time-slice within the constraints of limited storage.

\vspace{0.8em}

For example, imagine a space containing 10,000 objects, of which 10 changed state one second ago. We need only record the changes for those 10 objects at that moment; the states of the remaining 9,990 objects are simply inherited from the previous second's record. This process can be traced backwards continuously---if a given object has never changed, a single data entry suffices to describe its state at any point in time.

\newpage
\subsubsection{Is the Future Predictable?}
Since, in the Fifth Dimension, we can deterministically trace back the past, can we also look forward along the river of time to predict the future?

\vspace{1em}

Let us first revisit a classic episode from the history of physics, and see what we can learn from it.
After Newton published the law of universal gravitation, science entered a glorious golden age. From the orbits of celestial bodies to the combinations of molecules, everything could be expressed with exquisitely concise equations, accurate to the last decimal.

\vspace{1em}

``God does not play dice'' is Einstein's famous dictum---if you know, with perfect precision, the angular momentum of the throw, the friction coefficient of the die, its weight distribution, even the rotational speed of the Earth, then the number on the die is predictable.
When a prediction turns out to be wrong, classical determinism typically attributes the error to this: the environmental factors fed into the calculation were not yet complete enough.
Taken to its logical conclusion, the physical world seems as though, from the very instant of the Big Bang, \textbf{every subsequent trajectory was already written into the initial conditions.}
\vspace{1em}

Yet at the turn of the last century, two ``small clouds''\footnote{These two ``small clouds'' refer to the difficulties encountered with the luminiferous ether theory and the Maxwell--Boltzmann doctrine of equipartition of energy.} drifting across the apparently flawless, sunny sky of classical physics ultimately toppled its grand edifice, and upon its ruins was erected quantum physics. The famous ``simultaneously dead and alive'' cat of Schr\"{o}dinger from the history of quantum physics tells us: before the final observation, the state of an object exists in probabilistic form and cannot be determined with certainty.

\vspace{1em}

I have neither the intention nor the ability to discuss the still-unresolved thought experiments of quantum physics here, but I would like to share one insight with the reader: in the Fifth Dimension, many predictions can likewise exist only in probabilistic form. We can rarely arrive at fully deterministic inferences the way idealized classical equations do. As a general rule, systems with fewer variables and shorter time horizons are more likely to yield stable predictions.

\vspace{1em}

There is a Chinese proverb: \textbf{``A hair's breadth of error leads to a thousand miles astray.''} The more factors a prediction involves, the greater the probability of error,
and accumulated errors compound multiplicatively until the prediction ends up a thousand miles off course.\footnote{These ``thousand miles astray'' outcomes often open an entirely new door for the world.}
The black-and-white, deterministic thinking of classical physics has a deep-rooted hold on us. But from a Bayesian perspective, our understanding of the world does not arrive at truth in a single leap; instead, we set out with some prior understanding, and as new evidence arrives, we continuously revise our judgment.

\textbf{What} we call certainty is, much of the time, not an answer bestowed upon us from the start, but a high-confidence conclusion formed after repeated updates of our priors. In the Fifth Dimension, we need to learn to live with uncertainty. Just as humans are born with curiosity and the spirit of exploration, when we design algorithms, we should also leave room for those vanishingly improbable cases, edge states, and scenarios that have never yet occurred. \textbf{Do not view the world through the lens of certainty.} Allow low-probability events to exist. Allow models to be revised. Only then can our cognitive frameworks possess greater resilience, and only then can algorithms exhibit greater robustness in the real world---better equipped to face an ever-changing future.

\subsection{Spell Five: The Star-Absorbing Skill---Knowing Everything Means Knowing Nothing}
When I was a boy reading martial arts novels, I was especially envious of the master who could press his palm to his disciple's back and transmit decades of internal power in a single touch, transforming the student overnight. And Linghu Chong, with his Star-Absorbing Skill, accumulated immense reserves of inner strength that saved him from peril time and again.

\vspace{1em}

In the four-dimensional world, we are familiar with the 10,000-hour rule: to master a skill at an expert level, a human being must invest roughly ten thousand hours. I often tell my family that our children do not really need to study English deliberately---they just need to master their native Chinese. By the time they grow up, automatic translation will be a mature technology, capable of instantly translating any language on Earth.

After entering the Fifth Dimension, aided by artificial intelligence algorithms, abilities that once required humans to invest enormous amounts of time---translation, playing Go, identifying pulmonary nodules on CT scans---are increasingly being packaged as callable modules. These abilities, too, can be replicated through the Star-Absorbing Skill: once stably implemented, they can typically be invoked again and again (the loading process is usually instantaneous, far short of ten thousand hours), rapidly expanding the boundaries of what we can do.

\vspace{1em}

Once we have acquired these abilities, we can devote our finite time to more valuable pursuits. Each new capability module is stored in an arsenal of skills; the more skills in the arsenal, the greater one's overall strength, and the more time is saved. Time is life. In this sense, the Star-Absorbing Skill extends our effective lifespan.

\vspace{1em}

Another important change is that in the Fifth Dimension, memories and files can be preserved at low cost and for long periods. Through indexing technology, knowledge retrieval can be made nearly instantaneous---replacing the memorization of knowledge itself with the indexing of memory, thereby vastly expanding an individual's knowledge capacity.

\vspace{1em}

\textbf{More} and more people will come to possess a magic pocket like Doraemon's, ready at a moment's notice to enter the Fifth Dimension. Therefore, we no longer need to cram every piece of knowledge into our heads, block by block. Instead, we must learn to build an index for our knowledge. Whoever is better at invoking the index will be better at commanding knowledge.

\subsection{Spell Six: The Sinew Transformation Classic---Becoming a True Super-Individual}

In Linghu Chong's story, The Sinew Transformation Classic\footnote{If the previous spells have all been teaching us how to explore outward, this final internal discipline---The Sinew Transformation Classic---was added when I revisited the Fifth Dimension after large language models swept the globe. It will teach us how to cultivate inward.} resolved the dangerous aftereffects of the Star-Absorbing Skill, freeing him from the torment of conflicting internal energies and allowing him to harness the power of all martial arts for his own use, ultimately achieving legend.

\vspace{1em}

In the Fifth Dimension, we will put our learning into practice, using the methods of model training to train ourselves. We will treat our own private context as training data and our unique taste as the navigation system for that training, becoming the kind of person who stands \textbf{outside the distribution} of the era's average---and thereby becoming irreplaceable.

\vspace{1em}

As we practice this Sinew Transformation Classic, we ourselves will gradually become a personalized model: highly aligned, rapidly iterating, with an independent objective function.
During training, we must deliberately create resistance against the frictionless ease that AI offers us.

\vspace{1em}

When AI can produce mediocre-but-correct copy for us,
we must have the ability to see through that ``average-level perfection''
and, for the sake of that 5\% aesthetic gradient, be willing to tear it down and redo it dozens of times.
Only a private reward model forged in this way
gives us any chance of finding high ground when the algorithmic flood comes, rather than being swept under.

\vspace{1em}

The competition of the future is a competition of ``signal-to-noise ratio.'' The strength of the super-individual lies not in possessing a more powerful ``pump,'' but in possessing a more precise ``filter.'' When everything in the world can be Ctrl+Z'd, those who are willing to bear the irreversible cost of time---to accumulate private context, to calibrate their aesthetic gradient---are, in effect, using a kind of stubborn persistence to hedge against the certainty of algorithms.

\vspace{1em}

The true Sinew Transformation Classic is this: in a world made of bits, hold on to what is clumsy and human in us, and become a person unlike any other. Once we have our own objective function, AI is no longer a terrifying flood that will replace us, but the sword in our hand---the ``Dugu Nine Swords,'' where having no fixed technique surpasses all techniques.

The Star-Absorbing Skill lets us draw upon the power of heaven and earth. \textbf{The Sinew Transformation Classic makes us the master of those powers.}
\newpage

\section{Epilogue: The Laws of the Fifth Dimension}

\begin{myquote}
    Arthur C. Clarke said: ``Any sufficiently advanced technology is indistinguishable from magic.''
\end{myquote}

We possess these magical abilities because we have crossed a dimensional threshold. These six spells rest upon three fundamental laws of the Fifth Dimension. The spells show us what we can do in this dimension; the laws remind those of us who dwell within it how to judge and how to act.

\vspace{1em}

\phantomsection
\addcontentsline{toc}{subsection}{Law One: From Hoarding to Connecting---The Collapse of the Space-Time Barrier}
\textbf{Law One: From Hoarding to Connecting---The Collapse of the Space-Time Barrier}

In the past, security came from hoarding. Because the world was full of scarcity, people believed that only what was clutched in their hands was truly theirs, and they saw the movement of assets as a loss of protection. But in the Fifth Dimension, isolated data is dead; only bits that are plugged into a network and flowing continuously possess vitality. Connection is the only defense against marginalization---in this dimension, being isolated is more dangerous than being surveilled.

\vspace{0.8em}
\phantomsection
\addcontentsline{toc}{subsection}{Law Two: From Owning to Invoking---A Dimensional Shift in How Wealth Is Measured}
\textbf{Law Two: From Owning to Invoking---A Dimensional Shift in How Wealth Is Measured}

In the past, wealth followed a logic of stock: we defined identity by what we owned, prizing factories, land, and patent portfolios, with legal ownership determining who you were. In the Fifth Dimension, the truly powerful no longer chase legal title to vast server farms or raw data; they pursue the practical ability to marshal the world's computing power and data. Whether you are an individual who ``owns nothing,'' so long as you can ``invoke everything,'' you truly possess this world.

\vspace{0.8em}

\phantomsection
\addcontentsline{toc}{subsection}{Law Three: From Solving to Asking---The Great Abundance of Intelligence}
\textbf{Law Three: From Solving to Asking---The Great Abundance of Intelligence}

In the past, intelligence was tethered to the body, reset to zero at death, and experience was an exclusionary moat. But in the Fifth Dimension, the ability to solve problems is no longer scarce; what has become scarce is the intuition to define them. AI can absorb thousands of years of human knowledge overnight and load it on demand for whoever needs it. Intelligence is becoming the cheapest commodity on the industrial line; the gap between the veteran and the novice is flattened to zero in its presence. \textbf{While AI raises the floor (the lower bound) for everyone, it also raises the ceiling (the upper bound) between one person and another.}

\vspace{0.8em}

This is the eve of the storm we face: a new world whose laws are still being rewritten. Space and time no longer isolate us. Ownership no longer protects us. Experience no longer defines us. Since the tickets from the old world have been voided, we must build a new ark under the new laws of this world.

% 重置章节计数器，确保下一章是第一章
\setcounter{chapter}{0}
