\newpage
\section{Acknowledgements}
This book itself can be seen as an exercise in being a super individual. Through the collaboration of a human mind and AI, it went from a notion in my head to a published volume in remarkably short time. A thousand people have access to the same AI, yet a thousand people can write a thousand different Hamlets.
\vspace{1em}

The prologue of this book was written in the winter of 2020. My ambition at the time was to write a sequel to \emph{Being Digital}, introducing a worldview for the data age. But after finishing the prologue, I realized my knowledge was nowhere near sufficient to support that vision, so I set it aside.

\vspace{1em}

Five years passed. Along the entrepreneurial road I made many friends and lived through my share of joy and sorrow.
The crisis at the end of 2025 tormented me deeply. The pain was so intense that ``heartbreak'' was no longer just a metaphor---it became a physiological fact. Yet it was precisely that immense pain that compelled me, once the worst had passed, to pick up my pen again and continue this book.

\vspace{1em}

\textbf{Many people made this book possible.}

Thank you to my family, and to every reader. The support of my family and the feedback from readers gave me tremendous encouragement, confirming that this book truly can offer cognitive value to the younger generation.

\vspace{1em}

Thank you to my boss, Chen Tianqiao. He provided an environment free of worry, allowing me to examine the world's changes from a broad and forward-looking perspective. Pursuing deep data rather than big data has always been his expectation of us.
\vspace{1em}

Thank you to Celine. Many of the original ideas in this book grew out of discussions with her;
in a sense, she is a co-author.
Her product philosophy---centered on helping people---has deeply influenced me as well.

\vspace{1em}

That philosophy gradually clarified the book's purpose as writing continued---to write a book for the children and parents of this era who feel lost when facing career choices, to write an AI supplementary reader for high school students or college freshmen. In my view, truly \textbf{self-evolving AI} may still require a major architectural or learning-paradigm breakthrough. By the time these young readers enter the workforce, the world will have changed enormously---but they still have time to understand and adapt. Helping them better grasp the worldview of this era is, I believe, one thing this book can do for them.

\newpage
Ms.\ Dong Ying is the editor-in-chief of this book. Initially we agreed on a February 28 deadline (a seemingly impossible one). Throughout the process she continuously provided market intelligence and editorial feedback, helping this book reach more readers and nurturing a seed of thought into a real book\footnote{I later learned that the professional term for an editor helping an author shape a manuscript into a finished book is ``field management''---tending the crop from seedling to harvest.}.

\vspace{1em}

Xu Chao told me: ``Don't get lost in the technology. Technology is ultimately meant to serve people. Everyone can live a better life with AI's help---that's the aspiration of every ordinary person. Let me connect you with the education board officials.''
\vspace{1em}

Kuang Cuibiao said: ``I have an editor friend I've known for years, very reliable. Let me introduce you.''
\vspace{1em}

Charles said: ``Let me connect you with the Foreign Languages Press. This book deserves an English edition.''
\vspace{1em}

Zhang Zhijun encouraged me: ``Write a book that stands the test of time. Really, a person doesn't need to do many things in a lifetime. Writing one book is enough to justify having come into this world.''
\vspace{1em}

Wei Jia left a message: ``Mr.\ Jiang, after your book is published, could I ask you to sign two copies for me to give to children of ethnic minorities? If cost is no concern, I'd also like to send some to children in impoverished areas. Next year I plan to go to L\"{u}liang in Shanxi to help with education, and I hope to bring your book to them as well.''
\vspace{1em}

Along the way, many other friends---some I know well, some I have never met---offered me tremendous help. I cannot thank each of them here by name, but my gratitude is no less for it.
\vspace{1em}

I have set aside three thousand copies, hoping to donate them to education boards across the country as supplementary AI reading material for high school students. Three thousand is a modest number, but I have always believed that a book opened at the right moment can change the way a person sees the world. This book is written for young people, and for everyone living through this great transformation. If it can be a spark that ignites curiosity and courage about the future in the heart of some young person, that would be my greatest wish as an author.
